Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-28 paper/paper.tex

% \subsubsection{The Bellman equation:  characterizing optimal behavior}
% Q-learning (Watkins \& Dayan, 1992) generates a
% \textit{greedy policy}
\subsubsection{Optimal behavior and the Bellman equation}

Optimal behavior of the agent corresponds to
a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. Formally,
% \textbf{XXX please add 1-2 sentences why this is interesting in the current
% context. This statement appears a little alone as it stands.}
\begin{equation}
  \pi^*(s) := \argmax_{a \in \mathcal A}{Q}^*(s, a),\text{ where }Q^*(s, a) := \max_{\pi}Q^\pi(s,a).
  \label{eq:qlearning}
\end{equation}
The learning goal is to approach the
policy $\pi^*$ as ``close as possible'', that is to \text{solve} the MDP.
Of course \eqref{eq:qlearning} is just a definition and
represents by no means, a candidate schema for solving MDPs with even moderately-sized action and state spaces: it's intractable!
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without querying the entire space of policies, with the form
\begin{equation}
  Q^* = \text{Bel}(Q^*),
  \label{eq:bellman}
\end{equation}
where the ``Bellman transform'' $\text{Bel}(Q)$ of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is another $Q$-value function defined by
\begin{equation}
  \begin{split}
   \text{Bel}(Q)(s,a) &:=
   \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q(s', a')]\\
   &= r(s,a) + \gamma\mathbb E_{s' \sim p(s'|s,a)} [\max_{a' \in \mathcal A}Q(s', a')]\\
   &= \text{instantaneous reward} + \text{expected reward for acting greedily thereafter}
%    \; \forall (s,a) \in \mathcal S \times \mathcal A
    % \\&=
    % \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^*(s', \pi^*(s'))].
    \end{split}
  \end{equation}
The Bellman equation \eqref{eq:bellman} is a ``self-consistency'' equation which provides
a recursive decomposition of optimal behavior by dividing the $Q$-value function into the immediate
reward and the discounted rewards of the upcoming states in the environment.
The optimal Q-value operator $Q^*$ is a fixed point for this equation.
As a consequence of this decomposition, the complicated dynamic programming optimization
problem \eqref{eq:qlearning}
is broken into simpler sub-problems at different time points.
Exploitation of hierarchical structure in action considerations
has previously been related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value
to guide action towards a preferred state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Note that in \eqref{eq:bellman} the random sampling
is performed only over quantities which
depend on the environment.
This aspect of the learning process
can unroll off-policy by observing state transitions
triggered by another behavioral policy, which can be stochastic.

@@

>>>>

% \subsubsection{The Bellman equation:  characterizing optimal behavior}
% Q-learning (Watkins \& Dayan, 1992) generates a
% \textit{greedy policy}
\subsubsection{Optimal behavior and the Bellman temporal consistency}

Optimal behavior of the agent corresponds to
a strategy $\pi^*$ for choosing actions such that, for every state, the chosen action guarantees the best possible reward on average. Formally,
% \textbf{XXX please add 1-2 sentences why this is interesting in the current
% context. This statement appears a little alone as it stands.}
\begin{equation}
  \pi^*(s) := \argmax_{a \in \mathcal A}{Q}^*(s, a),\text{ where }Q^*(s, a) := \max_{\pi}Q^\pi(s,a).
  \label{eq:qlearning}
\end{equation}
The learning goal is to approach the
policy $\pi^*$ as ``close as possible'', that is to \text{solve} the MDP.
Of course \eqref{eq:qlearning} is just a definition and
represents by no means, a candidate schema for solving MDPs with even moderately-sized action and state spaces: it's intractable!
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
Fortunately, the \textit{Bellman equation} \citep{sutton1998reinforcement} provides a fixed-point relation which defines $Q^*$ implicitly via a sampling procedure, without querying the entire space of policies, with the form
\begin{equation}
  Q^* = \text{Bel}(Q^*),
  \label{eq:bellman}
\end{equation}
where the ``Bellman transform'' $\text{Bel}(Q)$ of an arbitrary $Q$-value function $Q: \mathcal S \times \mathcal A \rightarrow \mathbb R$  is another $Q$-value function defined by
\begin{equation}
  \begin{split}
   \text{Bel}(Q)(s,a) &:=
   \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q(s', a')]\\
   &= r(s,a) + \gamma\mathbb E_{s' \sim p(s'|s,a)} [\max_{a' \in \mathcal A}Q(s', a')]\\
   &= \text{instantaneous reward} + \text{expected reward for acting greedily thereafter}
%    \; \forall (s,a) \in \mathcal S \times \mathcal A
    % \\&=
    % \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^*(s', \pi^*(s'))].
    \end{split}
  \end{equation}
The Bellman equation \eqref{eq:bellman} is a temporal consistency equation which provides
a dynamic decomposition of optimal behavior by dividing the $Q$-value function into the immediate
reward and the discounted rewards of the upcoming states.
The optimal Q-value operator $Q^*$ is a fixed point for this equation.
As a consequence of this decomposition, the complicated dynamic programming optimization
problem \eqref{eq:qlearning}
is broken into simpler sub-problems at different time points.
Exploitation of hierarchical structure in action considerations
has previously been related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value
to guide action towards a preferred state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Note that in \eqref{eq:bellman} the random sampling
is performed only over quantities which
depend on the environment.
This aspect of the learning process
can unroll off-policy by observing state transitions
triggered by another behavioral policy, which can be stochastic.

@@

% \subsubsection{Bounded rationality}
% $$\max_{a \in \mathcal A}\tilde{Q}(s, a|\theta)  \propto \lim_{T \rightarrow 0^+}\exp(\tilde{Q}(s,a|\theta)/T)$$
